---
name: Tricia
surname: Van Laar
position: "Assistant Professor of Microbial Genomics"
profilepic: "img/tricia.jpg"
address: "Department of Biological Sciences, California State University Stanislaus"
phone: +1 209 667 3695
email: "tvanlaar@csustan.edu"
#twitter: tvanlaarmicro
github: tvanlaar
linkedin: tvanlaar
orcid: 0000-0002-4654-8501
date: "`r format(Sys.time(), '%B %Y')`"
#aboutme: "Dr. Van Laar is an educator and scientist interested in all things microbial."

headcolor: 414141
output: 
  vitae::awesomecv:
    page_total: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
library(here)
library(tidyverse)
```

# Education
```{r education}
tribble(
  ~ degree, ~ uni, ~ loc, ~ dates, ~ details,
  "MBA", "California State University, Stanislaus", "Fresno, CA", "Expected 2024", 
    list("Course work in Human Resources, New Venture Management, and Strategic Marketing"),
  "Ph.D. in Biology with an emphasis in Cell and Molecular Biology", "The University of Texas at San Antonio", "San Antonio, TX", "2012", 
    list("Dissertation Project: Role of the mevalonate pathway in the patho-physiology of \\textit{Borrelia burgdorferi}",
         "Major Advisor: J. Seshu, Ph.D."),
  "MS in Biology", "University of the Pacific", "Stockton, CA", "2007", 
    list("Thesis Project: The behavior of Rad51D and XRCC2 in response to drug induced DNA damage and a continuing study of the fly Rad51 paralogs",
         "Major Advisor: Lisa Wrischnik, Ph.D."),
   "BS in Biology", "California State University, Stanislaus", "Turlock, CA", "2005", 
    list("Minor in Chemistry, Concentration in Genetics",
         "Research Project: Speciation of \\textit{Sylvilagus bachmani riparius}",
         "Major Advisor: Jim Youngblom, Ph.D.")
) %>% 
  detailed_entries(degree, dates, uni, loc, details)
```

# Professional Positions
```{r professional}
tribble(
  ~ role, ~ company, ~ loc, ~ dates,
  "Assistant Professor of Microbial Genomics", "California State University, Stanislaus", "Turlock, CA", "08/2022 - Present", 
  "Associate Professor of Microbiology", "California State University, Fresno", "Fresno, CA", "08/2021-08/2022",
   "Assistant Professor of Microbiology", "California State University, Fresno", "Fresno, CA", "08/2015-08/2021",
  "Adjunct Instructor", "St. Philip's College", "San Antonio, TX", "08/2014-08/2015",
  "NRC Post-doctoral Fellow", "US Army Institute for Surgical Research, Dental and Trauma Research Detachment", "JBSA Ft. Sam Houston, TX", "04/2012-07/2015",
 "Graduate Research and Teaching Assistant", "The University of Texas at San Antonio", "San Antonio, TX", "08/2007-03/2012",
 "Graduate Research and Teaching Assistant", "University of the Pacific", "Stockton, CA", "08/2005-08/2005",
 "Undergraduate Research Assistant", "Endangered Species Recovery Program", "Turlock, CA", "04/2004-05/2005"
) %>% 
  detailed_entries(role, dates, company, loc)
```

<!-- # Publications Section -->

# Peer-Reviewed Publications (Student Authors Underlined)

\setlength{\leftskip}{0cm}
**2023**

\setlength{\leftskip}{1cm}
Whittaker, D. J., Atyam, A., Burroughs, N. A., Greenberg, J. M., Hagey, T. J., Novotny, M. V., Soini, H. A., Theis, K. R., **Van Laar, T. A.**, & Slade, J. W. G. (2023). Effects of short-term experimental manipulation of captive social environment on uropygial gland microbiome and preen oil volatile composition. Frontiers in Ecology and Evolution, 10. https://www.frontiersin.org/articles/10.3389/fevo.2022.1027399

\newpage 

\setlength{\leftskip}{0cm}
**2022**

\setlength{\leftskip}{1cm}
**Van Laar, T. A.**, \underline{Biehler, L.}, & Slade, J. W. G. (2022). Draft Genome Sequence of a Multidrug-Resistant Strain of *Salmonella enterica* Serovar Typhimurium Isolated from a Pine Siskin (*Spinus pinus*). Microbiology Resource Announcements, e0098222. https://doi.org/10.1128/mra.00982-22

\setlength{\leftskip}{0cm}
**2020**

\setlength{\leftskip}{1cm}
\underline{Osborne, C. J.}, \underline{Wakeman-Hill, A. J.}, \underline{Loa, S. E.}, Crosbie, P. R., & **Van Laar, T. A.** (2020). *Rickettsia* spp. In Five Tick Species Collected in Central California. Journal of Medical Entomology, 57(5), 1596–1603. https://doi.org/10.1093/jme/tjaa080

\underline{Avila, A. T.}, **Van Laar, T. A.**, Constable, J. V. H., & Waselkov, K. (2020). 16S rRNA Gene Diversity of Bacterial Endophytes in Parasitic *Cuscuta campestris* and Its *Helianthus annuus* Host. Microbiology Resource Announcements, 9(43). https://doi.org/10.1128/MRA.00968-20

\setlength{\leftskip}{0cm}
**2019**

\setlength{\leftskip}{1cm}
\underline{Osborne, C. J.}, Crosbie, P. R., & **Van Laar, T. A.** (2019). *Borrelia parkeri* in *Ornithodoros parkeri* (Ixodida: Argasidae) Collected Using Compact Dry Ice Traps in Madera County, California. Journal of Medical Entomology, 56(2), 579–583. https://doi.org/10.1093/jme/tjy213

\underline{Esani, S.}, Chen, T., Leung, K. P., & **Van Laar, T. A.** (2019). Transcriptome Sequence of Antibiotic-Treated *Pseudomonas aeruginosa*. Microbiology Resource Announcements, 8(12). https://doi.org/10.1128/MRA.01367-18

\setlength{\leftskip}{0cm}
**2018**

\setlength{\leftskip}{1cm}
**Van Laar, T. A.**, \underline{Esani, S.}, \underline{Birges, T. J.}, Hazen, B., \underline{Thomas, J. M.}, & Rawat, M. (2018). *Pseudomonas aeruginosa gshA* Mutant Is Defective in Biofilm Formation, Swarming, and Pyocyanin Production. mSphere, 3(2). https://doi.org/10.1128/mSphere.00155-18

\setlength{\leftskip}{0cm}
**2017**

\setlength{\leftskip}{1cm}
Miller, C. L., **Van Laar, T. A.**, Chen, T., Karna, S. L. R., Chen, P., You, T., & Leung, K. P. (2017). Global transcriptome responses including small RNAs during mixed-species interactions with methicillin-resistant *Staphylococcus aureus* and *Pseudomonas aeruginosa*. MicrobiologyOpen, 6(3). https://doi.org/10.1002/mbo3.427

Lin, Y.-H., Romo, J. A., Smith, T. C. 2nd, Reyes, A. N., Karna, S. L. R., Miller, C. L., **Van Laar, T. A.**, Yendapally, R., Chambers, J. P., & Seshu, J. (2017). Spermine and Spermidine Alter Gene Expression and Antigenic Profile of *Borrelia burgdorferi*. Infection and Immunity, 85(3). https://doi.org/10.1128/IAI.00684-16

\setlength{\leftskip}{0cm}
**2016**

\setlength{\leftskip}{1cm}
**Van Laar, T. A.**, Hole, C., Rajasekhar Karna, S. L., Miller, C. L., Reddick, R., Wormley, F. L., & Seshu, J. (2016). Statins reduce spirochetal burden and modulate immune responses in the C3H/HeN mouse model of Lyme disease. Microbes and Infection, 18(6), 430–435. https://doi.org/10.1016/j.micinf.2016.03.004

\setlength{\leftskip}{0cm}
**2015**

\setlength{\leftskip}{1cm}
**Van Laar, T. A.**, Chen, T., You, T., & Leung, K. P. (2015). Sublethal concentrations of carbapenems alter cell morphology and genomic expression of *Klebsiella pneumoniae* biofilms. Antimicrobial Agents and Chemotherapy, 59(3), 1707–1717. https://doi.org/10.1128/AAC.04581-14

\setlength{\leftskip}{0cm}
**2013**

\setlength{\leftskip}{1cm}
Childers, B. M., **Van Laar, T. A.**, You, T., Clegg, S., & Leung, K. P. (2013). MrkD1P from *Klebsiella pneumoniae* strain IA565 allows for coexistence with *Pseudomonas aeruginosa* and protection from protease-mediated biofilm detachment. Infection and Immunity, 81(11), 4112–4120. https://doi.org/10.1128/IAI.00521-13

\setlength{\leftskip}{0cm}
**2012**

\setlength{\leftskip}{1cm}
**Van Laar, T. A.**, Lin, Y.-H., Miller, C. L., Karna, S. L. R., Chambers, J. P., & Seshu, J. (2012). Effect of levels of acetate on the mevalonate pathway of *Borrelia burgdorferi*. PloS One, 7(5), e38171. https://doi.org/10.1371/journal.pone.0038171
\newpage
\setlength{\leftskip}{0cm}
**2011**

\setlength{\leftskip}{1cm}
Raju, B. V. S., Esteve-Gassent, M. D., Karna, S. L. R., Miller, C. L., **Van Laar, T. A.**, & Seshu, J. (2011). Oligopeptide permease A5 modulates vertebrate host-specific adaptation of *Borrelia burgdorferi*. Infection and Immunity, 79(8), 3407–3420. https://doi.org/10.1128/IAI.05234-11

\setlength{\leftskip}{0cm}

# Editor Accepted Publications (Student Authors Underlined)

**2017**

\setlength{\leftskip}{1cm}
\underline{Esani, S.}, Constable, J. V. H., & **Van Laar, T. A.** (2017). Draft Genome Sequence of a Multidrug-Resistant Strain of *Enterococcus faecalis*, PM01, Isolated from the Nest of an American Bushtit, *Psaltriparus minimus*. Genome Announcements, 5(11). https://doi.org/10.1128/genomeA.00017-17

\setlength{\leftskip}{0cm}
**2016**

\setlength{\leftskip}{1cm}
\underline{Kennedy, V.}, **Van Laar, T. A.**, \underline{Aleru, O.}, Thomas, M., Ganci, M., & Rawat, M. (2016). Genome Sequences of Three Spore-Forming Bacteria Isolated from the Feces of Organically Raised Chickens. Genome Announcements, 4(5). https://doi.org/10.1128/genomeA.00880-16

\underline{Nelson, R. L.}, \underline{Castro, M. A.}, Katti, M., Eisen, J. A., & **Van Laar, T. A.** (2016). Genome Sequence of a Multidrug-Resistant Strain of *Bacillus pumilus*, CB01, Isolated from the Feces of an American Crow, *Corvus brachyrhynchos*. Genome Announcements, 4(4). https://doi.org/10.1128/genomeA.00807-16

\setlength{\leftskip}{0cm}
**2014**

\setlength{\leftskip}{1cm}
**Van Laar, T. A.**, Chen, T., Childers, B. M., Chen, P., Abercrombie, J. J., & Leung, K. P. (2014). Genome Sequence of a Multidrug-Resistant Strain of *Klebsiella pneumoniae*, BAMC 07-18, Isolated from a Combat Injury Wound. Genome Announcements, 2(6). https://doi.org/10.1128/genomeA.01230-14

\setlength{\leftskip}{0cm}

# Book Chapters

Seshu, J., Smith, T. C. 2nd, Lin, Y.-H., Karna, S. L. R., Miller, C. L., & **Van Laar, T. A.** (2018). Analysis of DNA and RNA Binding Properties of *Borrelia burgdorferi* Regulatory Proteins. Methods in Molecular Biology (Clifton, N.J.), 1690, 155–175. https://doi.org/10.1007/978-1-4939-7383-5_13

# Invited Seminars

```{r seminars}
tribble(
  ~ title, ~ organization, ~ location, ~ year,
  "Math is for the Birds", "Seminar Series for the Quanititatively Curious", "Stanislaus State", "2023",
  "Microbiology of Avian Systems", "Seminar Series", "University of the Pacific", "2023",
   "Avian Microbial Genomics", "Seminar Series", "Stanislaus State", "2023",
  "The Birds of Fresno State", "Fresno Audubon Society", "", "2020",
  "We did something bad: Is this the end of the antibiotic era?", "Colloquium", "Fresno State", "2019",
  "Good drugs gone bad: Antimicrobial resistance research at Fresno State", "Seminar Series", "UC Merced", "2018",
  "GMOs and transgenic organisms – threat or opportunity?", "Central Valley Cafe Scientifique", "Fresno, CA", "2018",
  "Good drugs gone bad: Antimicrobial Resistance", "Central Valley Cafe Scientifique", "Fresno, CA", "2017",
  "Good drugs gone bad: An investigation of antimicrobial/resistance and tolerance in war-wound pathogens", "Seminar Series", "Stanislaus State", "2016",
  "A biologist and a sociologist walk into a…CRISPR", "Central Valley Cafe Scientifique", "Fresno, CA", "2015",
  "Role of the mevalonate pathway in the patho-physiology of Borrelia burgdorferi", "South Texas Center for Emerging Infectious Diseases", "The University of Texas at San Antonio", "2011",
  
) %>% 
  brief_entries(glue::glue("\\textbf{<title>}: <organization>, <location>", .open = "<", .close = ">"), paste0(year), .protect = FALSE)
```

# Grants and Funding (Total funding: $804,891)

Stanislaus State
```{r}
tribble(
  ~ award, ~ year, ~ amount,
  "Research, Scholarly, and Creative Activity Award", "2023", "10,000",
  "National Science Foundation New Faculty Enhancement Grant (Co-PI)", "2022", "467,622",
  "Research, Scholarly, and Creative Activity Award", "2022", "10,000",
  
) %>%
brief_entries(glue::glue("\\textbf{{<award>}} \\hfill \\${{<amount>}}", .open = "<", .close = ">"), paste0(year), .protect = FALSE)
```
\newpage
Fresno State
```{r}
tribble(
  ~ award, ~ year, ~ amount,
  "CSUPERB New Investigator Seed Grant (Co-PI)", "2022", "14,919", 
  "College of Science and Mathematics Scholarly and Creative Activity Award", "2020", "5,000",
  "CSU-STEM NET Grant (Co-PI)", "2019", "50,000",
  "CSU Program for Education and Research in Biotechnology Curriculum Redesign Award", "2019", "15,000",
  "College of Science and Mathematics Professional Development Award", "2018", "5,000",
  "College of Science and Mathematics Scholarly and Creative Activity Award (Co-PI)", "2018", "5,000",
  "College of Science and Mathematics Scholarly and Creative Activity Award (Co-PI)", "2018", "5,000",
  "Biology Travel Award", "2017", "1,000",
  "College of Science and Mathematics Scholarly and Creative Activity Award", "2017", "7,000",
  "College of Science and Mathematics Scholarly and Creative Activity Award", "2017", "5,000",
  "College of Science and Mathematics Professional Development Award", "2017", "950",
  "Biology Travel Award", "2016", "500",
  "College of Science and Mathematics Professional Development Award", "2016", "1,200",
  "Biology Travel Award", "2015", "1,000",
  
) %>%
  brief_entries(glue::glue("\\textbf{{<award>}} \\hfill \\${{<amount>}}", .open = "<", .close = ">"), paste0(year), .protect = FALSE)
```

Older Funding
```{r}
tribble(
  ~ award, ~ year, ~ amount,
  "National Research Council Post-Doctoral Fellowship", "2012", "150,000",
  "Center for Excellence in Infection Genomics Doctoral Scholarship", "2011", "30,000",
  "South Texas Center for Emerging Infectious Diseases PhD Fellowship", "2010", "30,000",
  "UTSA Presidential Dissertation Fellowship", "2010", "1,500"
  
) %>%
  brief_entries(glue::glue("\\textbf{{<award>}} \\hfill \\${{<amount>}}", .open = "<", .close = ">"), paste0(year), .protect = FALSE)
```

# Honors and Awards

Fresno State
```{r}
tribble(
  ~ award, ~ year,
  "Provost Award for Promising New Faculty", "2020",
  "Outstanding Advisor Nominee", "2019"
) %>%
  brief_entries(award, year)
```

Older Awards
```{r}
tribble(
  ~ award, ~ year,
  "O.B. Williams Award for Oral Presentation at Texas ASM meeting", "2009",
  "First Place Oral Presentation Stanislaus State Research Competition", "2005"
) %>%
  brief_entries(award, year)
```

<!-- # Outreach and Service Section -->
# Outreach and Service

Stanislaus State

*	Darwin Day Presenter (2024)
* CSUBIOTECH Nagel Award Selection Committee (2023-Present)
*	ASPIRE STEM 4 PALS Presenter (2023-present)
*	Social Media (Instagram/Twitter) (2022-Present)

Fresno State

*	American Society for Microbiology Abstract and Travel Award Reviewer (2020-present)
*	Fresno Audubon Society Campus Walks (2019-2022)
*	American Ornithological Society Poster Judge (2018-present)
*	College of Science and Mathematics Day of Giving (2018)
*	Feria de Educacion/Day of Discovery (2018-2022)
*	Science Olympiad Preparation with University High School (2017 & 2018)
*	Central California Research Symposium Judge (2017-2021)
*	Annual Biomedical Conference for Minority Students Presentation Judge (2016)
*	Café Scientifique Website, Facebook Page, Twitter Account (2016-2022)
*	Division of Graduate Studies Symposium Judge (2016)
*	Head Microbiology Judge for the Central California Regional Science, Math & Engineering Fair (2016)
*	Expanding Your Horizons Presenter (2015-2018)

\newpage
<!-- # Professional Society Membership Section -->
# Professional Society Membership

* American Ornithological Society
* American Society for Microbiology

<!-- # Professional Service Section -->
# Professional Service

* Associate editor for mSystems 
* Reviewer for Genes, Medical Sciences, mSystems, Microbiome, Microorganisms, Molecules, Molecular Biology and Evolution, Data, Biomolecules, Pharmaceuticals
* Reviewer for NSF 
* Reviewer for Microbiology: An Evolving Science 5e Textbook (Norton Publishing)

